A Recipe for Scavenging \textcolor{blue}{in Vertebrates} - the natural history of a behaviour

### 2016-07-29 natural history of scavenging.tex
Commented out some of Kev's red comments
They’re now using the %KH: format

\textcolor{blue}{While it is difficult to infer the metabolisms of extinct species, similarities can still also be drawn. Mainly ectothermic groups such as sharks and bony fish were likely to benefit from the ability to bridge long time-spans between meals such as seen in snakes. In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have at least partial endothermy resulting in requiring a higher intake of material in comparison to their ectothermic equivalents. However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.} \textcolor{red}{I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species}

@@

\textcolor{blue}{The physical differences between water and air mean carcass availability is radically different between these environments \citep{beasley2012carrion}. 
For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers.
Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers \citep{beasley2012carrion}. 
\textcolor{blue}{Research has shown that an animal need only travel 36 km to encounter a fresh whale carcass \citep{smith2003ecology}.} 
The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.}
\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
So, despite being unlikely \textcolor{red}{Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?}, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.

Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
This has a significant bearing on the availability and persistence of carrion.
We know from the geological record that the Earth has undergone radical fluctuations in temperature over time.
\textcolor{blue}{On land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
All of these factors influence the availabiltiy of carrion.  
Moreover, oceanic productivity is also impacted by climactic conditions. \textcolor{red}{I would delete the last two lines of this paragraph and merge it with the next one}

To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
The impacts these can have on scavengers have been empirically supported e.g. \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect} \textcolor{red}{This is a biotic interaction so I would either remove it or move it somewhere else}. 
The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.
Notably, freezing carcasses can become too hard to consume by most vertebrate carnviores \citep{selva2003scavenging}.}

>>>>

\textcolor{blue}{While it is difficult to infer the metabolisms of extinct species, similarities can still also be drawn. Mainly ectothermic groups such as sharks and bony fish were likely to benefit from the ability to bridge long time-spans between meals such as seen in snakes. In contrast, forms such as dinosaurs \citep{grady2014evidence}, pterosuars \citep{wellnhofer1991illustrated}, some marine reptiles \citep{bernard2010regulation} and some early mammals \citep{lovegrove2016phenology} have been proposed to have at least partial endothermy resulting in requiring a higher intake of material in comparison to their ectothermic equivalents. However, such higher metabolic rates would also increased the ability of such species to actively search for carrion.}
%KH: I put this section in as there is no reference in how we can use the above paragraph to infer scavenging in extinct species

@@

\textcolor{blue}{The physical differences between water and air mean carcass availability is radically different between these environments \citep{beasley2012carrion}. 
For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers.
Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers \citep{beasley2012carrion}. 
\textcolor{blue}{Research has shown that an animal need only travel 36 km to encounter a fresh whale carcass \citep{smith2003ecology}.} 
The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.}
\cite{ruxton2005searching} argued that although this is energetically feasible it's ecologically unlikely.
Any animal that could find such whale carcasses is unlikely to have ignored other types of carrion.
Although no aquatic species have ever exceeded the size of whales, some enormous animals have evolved in this environment before the evolution of cetaceans, including \textit{Leedsichthys}, a bony fish from the Middle Jurassic (174.1 - 163.5 Mya) and the aquatic Mesozoic reptiles, the plesiosaurs, pliosaurs and ichtyosaurs, that could all exceed 15 metres in length \citep{ruxton2011zoology,danise2014ecological}.
So, despite being unlikely 
% KH: Is there an estimate of how much of the fishes diet can be made up from this type of scavenging?
, the energetic feasibility of a marine scavenger that specialises on large carcasses has a long history.

Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
This has a significant bearing on the availability and persistence of carrion.
We know from the geological record that the Earth has undergone radical fluctuations in temperature over time.
\textcolor{blue}{On land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
All of these factors influence the availabiltiy of carrion.  
Moreover, oceanic productivity is also impacted by climactic conditions. 
%KH: I would delete the last two lines of this paragraph and merge it with the next one

To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
The impacts these can have on scavengers have been empirically supported e.g. \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}.
%KH: This is a biotic interaction so I would either remove it or move it somewhere else. 
The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.
Notably, freezing carcasses can become too hard to consume by most vertebrate carnviores \citep{selva2003scavenging}.}
